---
title: "RandomForest-Model"
author: "Dr. Simon Müller"
date: "`r Sys.Date()`"
output:
  html_vignette:
    css: kable.css
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{RandomForest-Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Initialization 

In the first example, we use the RandomForest based model and the `ovarian` data set from the 
`survival` package. In the first step we initialize the R6 data object. 

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(survival)
library(CaseBasedReasoning)
ovarian$resid.ds <- factor(ovarian$resid.ds)
ovarian$rx <- factor(ovarian$rx)
ovarian$ecog.ps <- factor(ovarian$ecog.ps)

# initialize R6 object
rfModel <- RFModel$new(Surv(futime, fustat) ~ age + resid.ds + rx + ecog.ps, data = ovarian, importance = "impurity_corrected")
```
All cases with missing values in the learning and end point variables are dropped (`na.omit`) and the reduced data set without missing values is saved internally. You get a text output on how many cases were dropped. `character` variables will be transformed to `factor`.

## Similar Cases 

After the initialization, we may want to get for each case in the query data the most similar case from the learning data. 
```{r}
n <- nrow(ovarian)
trainID <- sample(1:n, floor(0.8 * n), F)
testID <- (1:n)[-trainID]

# fit model 
ovarian[trainID, ] %>% 
  rfModel$fit()
# get similar cases
ovarian[trainID, ] %>%
  rfModel$get_similar_cases(query = ovarian[testID, ], k = 3) -> matchedData
knitr::kable(head(matchedData))
```

All cases with missing values in the learning and end point variables are dropped (`na.omit`) and the reduced data set without missing values is saved internally. You get a text output on how many cases were dropped. `character` variables will be transformed to `factor`.

Optionally, you may want to adjust some parameters in the fitting step of the random forest algorithm. Possible arguments are: , `ntree`, `mtry`, and `splitrule`. The documentation of this parameters can be found in the ranger R-package. Furthermore, you are able to choose the two distance measures:

+ proximity matrix 
+ depth matrix

This can be done by

```{r, warning=FALSE, message=FALSE}
rfModel$set_dist(distMethod = "proximity")
```
or
```{r, warning=FALSE, message=FALSE}
rfModel$set_dist(distMethod = "depth") # the default value
```

All other steps (excluding checking for proportional hazard assumption are the same as for the Cox-Model). 